Recent demography drives changes in linked selection across the maize genome

### 2015-11-24 NatGen/manuscript.tex
abstract

\begin{abstract}
  \noindent \bf
  \noindent
  The interaction between genetic drift and selection in shaping genetic diversity is not fully understood. In particular, a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$), but rapidly changing population demographics may complicate this relationship. To better understand how changing demography impacts selection, we investigated linked selection in the genomes of 23 domesticated maize and 13 wild maize (teosinte) individuals. We show that maize went through a domestication bottleneck with a population size of approximately 5\% that of teosinte before it experienced rapid expansion post-domestication. We observe that hard sweeps on genic mutations are not the primary force driving maize evolution. As expected, a reduced population size during domestication decreased the efficiency of purifying selection to purge deleterious alleles from maize, but rapid expansion after domestication has since increased the efficiency of purifying selection to levels exceeding those seen in teosinte. This final observation demonstrates that rapid demographic change can have wide-ranging impacts on diversity that conflict with would-be expectations based on long-term $N_e$.
\end{abstract}

>>>>

\begin{abstract}
  \noindent \bf
  \noindent
Across species, genetic diversity is shaped by the interaction of drift and selection. But the details of this interaction are not well understood. Although a population's propensity to drift is typically summarized by its long-term effective population size ($N_e$ ), rapidly changing population demographics complicate this relationship. To better understand how changing demography impacts selection, we used whole-genome sequencing data to investigate patterns of linked selection in domesticated and wild maize (teosinte). We produce the first unconfounded estimate of $N_e$ at the onset of maize domestication, which shows that maize was reduced to approximately 5\% the population size of teosinte before it experienced rapid expansion post-domestication. As expected, a reduced population size during domestication decreased the efficiency selection, but rapid expansion after domestication has since increased the efficiency of selection to levels exceeding those seen in teosinte. This final observation is applicable to all species that have undergone recent demographic change, including humans, domesticates, model, and non-model species. It demonstrates that rapid demographic change can have immediate and wide-ranging impacts on diversity that conflict with would-be expectations based on $N_e$ alone.
\end{abstract}
